---
title: "Basic Income Survey (part 1.)"
author: 'Anton Aksyonov'
date: 'may 2017'
output:
  html_document:
    number_sections: true
    toc: true
    fig_width: 9
    fig_height: 5
    theme: yeti
    highlight: textmate
---
# Intro
The current work was carried out to study the data and interpret the result of the study that was conducted by Dalia Research in April 2016

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(data.table)
library(ggplot2)
library(plotly)

income_data <- fread(paste0(getwd(),"/basic_income_dataset_dalia.csv"))
```

# Who are our respondents?
## Age, gender & education
The distribution of respondents by age among men and women is almost the same.
```{r, message=FALSE, warning=FALSE}
# DISTRIBUSION BY AGE + GENDER
ggplot(data = income_data, aes(x = age)) + 
  ggtitle('The distribution of respondents by age') +
  geom_histogram(aes(y = ..density..)
                 ,col = 'black'
                 ,fill = 'white') +
  geom_density(alpha = 0.2, fill='#FF6666') + 
  facet_grid(gender ~.)

```

Noticeably a slight shift towards men.
```{r, message=FALSE, warning=FALSE}
# DISTRIBUSION BY GENDER (SHARE)
ggplotly(
ggplot(data = income_data, aes(x = age, fill = gender)) + 
  ggtitle('Share distribution by male and female') +
  labs(x = 'Age', y = 'Share of gender') +
  geom_histogram(position = 'fill', alpha=0.7, binwidth = 1) + 
  scale_x_continuous(breaks=seq(10 , max(income_data[,age]), 5)) + 
  geom_hline(aes(yintercept = 0.5), colour="white")
)
```

How did the respondents share the presence of children?
```{r, message=FALSE, warning=FALSE}
ggplotly(
ggplot(data = income_data, aes(x = age, fill = dem_has_children)) + 
  ggtitle('Share distribution by presence of children  (male and female)') +
  labs(x = 'Age', y = 'Share presence of children') +
  geom_histogram(position = 'fill', alpha=0.7, binwidth = 1) + 
  scale_x_continuous(breaks=seq(10 , max(income_data[,age]), 5)) + 
  geom_hline(aes(yintercept = 0.5), colour="white") + 
  facet_grid(gender ~ .)
)
```


## What about education?

I changed the name of the attributes in the field `dem_education_level`, to rank these parameters.
```{r, message=FALSE, warning=FALSE, results='hide'}
income_data[dem_education_level == 'no', dem_education_level:="no"]
income_data[dem_education_level == 'low', dem_education_level:="3. low"]
income_data[dem_education_level == 'medium', dem_education_level:="2. medium"]
income_data[dem_education_level == 'high', dem_education_level:="1. high"]
```

Now let's see how our correspondents are distributed by age and level of education
```{r,, message=FALSE, warning=FALSE}
ggplotly(
  ggplot(data = income_data, aes(x = age, fill = dem_education_level)) + 
    ggtitle('Distribusion by age and level of education') + 
    labs(x = 'Number fo respondents', y = 'Share') + 
    geom_histogram(bins = 50
                 ,alpha = 0.7)
)
```
Let's look at the share of distribution. The majority of respondents have medium and higher education.
```{r, message=FALSE, warning=FALSE}
ggplotly(
  ggplot(data = income_data, aes(x = age, fill = dem_education_level)) + 
    ggtitle('Share of education level by age group') + 
    labs(x = 'Age Group', y = 'Share') + 
    geom_histogram(bins = 50
                   ,position = "fill"
                   ,alpha = 0.7)
)
```

Let's see how the share of the education level is distributed according to the age category of the respondents.
```{r, message=FALSE, warning=FALSE}
round(prop.table(table(income_data$dem_education_level, income_data$age_group),margin = 2),2)
temp_t <-  data.table(round(prop.table(table(income_data$dem_education_level, income_data$age_group),margin = 2),2))
names(temp_t) <- c('dem_education_level', 'age_group', 'prop')
```
Now let's display the shares on the bar-plot
```{r, message=FALSE, warning=FALSE}
ggplotly(
ggplot(data = temp_t, aes(x = age_group, y = prop, fill = dem_education_level)) + 
  ggtitle('Share of education level by age group') + 
  labs(x = 'Age Group', y = 'Share') + 
  geom_bar(stat = 'identity'
           ,alpha=0.7
           ,col = 'black')
)

```
Does the level of education affect the availability of regular work?

```{r,, message=FALSE, warning=FALSE}
round(prop.table(table(income_data$dem_education_level, income_data$dem_full_time_job),1),2)
temp_t <- data.table(round(prop.table(table(income_data$dem_education_level, income_data$dem_full_time_job),1),2))
names(temp_t) <- c('dem_education_level', 'dem_full_time_job', 'prop')
```

```{r,, message=FALSE, warning=FALSE}
ggplotly(
ggplot(data = temp_t, aes(x = dem_education_level, y = prop, fill = dem_full_time_job)) + 
  ggtitle('The proportion of respondents with a full time job,\n by education level') + 
  labs(x = 'Education level', y = 'Share') + 
  geom_bar(stat = 'identity'
           ,alpha=0.7
           ,col = 'black')
 
)
```
# Questions
For convenience, I will shorten field names with questions.
```{r, message=FALSE, warning=FALSE, results='hide'}
setnames(income_data
         , old = c(c('question_bbi_2016wave4_basicincome_awareness'
                     ,'question_bbi_2016wave4_basicincome_vote'
                     ,'question_bbi_2016wave4_basicincome_effect'
                     ,'question_bbi_2016wave4_basicincome_argumentsfor'
                     ,'question_bbi_2016wave4_basicincome_argumentsagainst'))
         , new = c(c('awareness'
                     ,'vote'
                     ,'effect'
                     ,'argumentsfor'
                     ,'argumentsagainst')))
# Numeric variable for aggregating the number of respondents
income_data$n <- 1
```

## How familiar are you with the concept known as “basic
income”? 
Prepare the data to analyze the results of the first question
```{r, message=FALSE, warning=FALSE, results='hide'}
temp_t_plot <- income_data[,.('num_ans' = sum(n)), by = c('awareness','gender', 'age_group', 'dem_education_level')]
unique(temp_t_plot$awareness)

temp_t_plot[awareness == "I understand it fully", awareness:=paste0('1.',"I understand it fully")]
temp_t_plot[awareness == "I know something about it", awareness:=paste0('2.',"I know something about it")]
temp_t_plot[awareness == "I have heard just a little about it", awareness:=paste0('3.',"I have heard just a little about it")]
temp_t_plot[awareness == "I have never heard of it", awareness:=paste0('4.',"I have never heard of it")]


```

```{r, message=FALSE, warning=FALSE}
ggplot(data = temp_t_plot, aes(x = dem_education_level, y = num_ans, fill = awareness)) + 
  ggtitle('How familiar are you with the concept known as “basic
income”?') +
  labs(x = 'Education level', y = 'Number of respondents') + 
  geom_bar(stat = 'identity') + facet_grid(gender ~ age_group) + 
  theme(axis.text.x = element_text(angle=45, vjust=0.5, size=8)) + 
  theme(legend.position = 'bottom'
        , legend.text = element_text(size = 7))
```

```{r, message=FALSE, warning=FALSE}
ggplot(data = temp_t_plot, aes(x = dem_education_level, y = num_ans, fill = awareness)) + 
  ggtitle('How familiar are you with the concept known as “basic
income”?') +
  labs(x = 'Education level', y = 'Share of respondents') + 
  geom_bar(stat = 'identity', position = 'fill') + facet_grid(gender ~ age_group) + 
  theme(axis.text.x = element_text(angle=45, vjust=0.5, size=8)) + 
  theme(legend.position = 'bottom'
        , legend.text = element_text(size = 7))
```
To be continue...
